### 2024-10-10 Code/MSB1015-Project.Rmd
Update MSB1015-Project.Rmd
Fixed some library issues, and divided the random forest chunk into multiple sub-chunks for easier use. Additionally, I added code to perform unsupervised random forest and visualised it using a function I made.

---
title: "MSB1015 Project"
author: "Carlo Alberto Zani"
date: "2024-09-08"
output: html_document
---

#0. Setup
```{r setup, include=FALSE}
library(rstudioapi)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(GGally)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()

# data <- read_csv("METABRIC_RNA_Mutation.csv")
```

# Data Cleaning
```{r cleaning, echo=FALSE}
# # Check for duplicates
# any(duplicated(data))
# 
# # Check for missing values
# missing_counts <- data %>%
#   summarise(across(everything(), ~ sum(is.na(.)))) %>%
#   pivot_longer(everything(), names_to = "column", values_to = "missing_count") %>%
#   filter(missing_count > 10)
# 
# # Plot where those missing values were found
# ggplot(missing_counts, aes(x = reorder(column, missing_count), y = missing_count)) +
#   geom_bar(stat = "identity", fill = "skyblue1") +
#   coord_flip() +
#   labs(x = "Column", y = "Missing Count", title = "Columns with More Than 10 Missing Values")
# 
# data$tumor_stage <- as.character(data$tumor_stage)
# 
# data$neoplasm_histologic_grade <- as.character(data$neoplasm_histologic_grade)
# 
# # Apply the modifications
# data_filled <- data %>%
#   
#   # 1. Replace missing values with "Unknown" for all categorical columns
#   mutate(oncotree_code = replace_na(oncotree_code, "Unknown"),
#          cancer_type_detailed = replace_na(cancer_type_detailed, "Unknown" )) %>%
#   
#   mutate(across(where(is.character), ~ factor(replace_na(., "Unknown")))) %>%
#   
#   # 3. Replace missing values with the median for "tumor_size"
#   mutate(tumor_size = ifelse(is.na(tumor_size), median(tumor_size, na.rm = TRUE), tumor_size)) %>%
#   
#   # 4. Replace missing values with 0 for columns ending with "_mut"
#   mutate(across(ends_with("_mut"), ~ replace_na(., 0))) %>%
#   
#   # 5. Convert the columns ending with "_mut" to a logical vector
#   mutate(across(ends_with("_mut"), ~ ifelse(. != "0", 1, 0))) %>%
#   
#   # 6.  
#   rowwise() %>%
#   mutate(mutation_count = sum(across(ends_with("_mut")) == 1))
# 
# # Check for missing values after cleaning
# sum(is.na(data_filled))
# 
# # Select only the transcription data
# pca_gx <- data_filled %>%
#   select(brca1:ugt2b7)
# 
# # Perform PCA to check the columns with the highest amount of missing values
# pca <- prcomp(pca_gx)
# 
# # Convert PCA results to data frame
# pca_data <- as.data.frame(pca$x[, 1:3])
# pca_data$tumor_stage <- as.factor(data_filled$tumor_stage)
# 
# # Create pairwise plots using ggpairs
# ggpairs(pca_data, mapping = aes(color = tumor_stage), 
#         upper = list(continuous = wrap("points")),
#         lower = list(continuous = wrap("points")),
#         diag = list(continuous = wrap("barDiag"))) +
#   theme_minimal()
```

# Imputation
```{r impute, echo=FALSE}

# # Load necessary libraries
# # library(mice)
# 
# # Get rid of unwanted characters from the column names
# colnames(data_filled) <- make.names(colnames(data_filled), unique = TRUE)
# 
# # Convert all character variables to factors
# data_filled <- data_filled %>%
#   mutate(across(where(is.character), as.factor))
# 
# # Wrap 'tumor_stage' as a factor
# data_filled$tumor_stage <- as.factor(data_filled$tumor_stage)
# 
# data_filled$neoplasm_histologic_grade <- as.factor(data_filled$neoplasm_histologic_grade)
# 
# # Identify categorical variables
# fac_vars <- names(data_filled)[sapply(data_filled, is.factor)]
# 
# # Specify methods for all variables, setting CART for categorical variables
# meth <- make.method(data_filled)
# meth[fac_vars] <- "cart" # Use CART for all categorical variables
# 
# # Perform the imputation using mice
# imputation_result <- mice(data_filled, method = meth, m = 5)
# 
# imputed_data <- complete(imputation_result)
# 
# write.csv(imputed_data, file = "imputed_data.csv", row.names = FALSE)
```

# Plots for imputation
```{r Plots after imputation}
# # Convert 'tumor_stage' to a factor in both datasets
# data$tumor_stage <- as.factor(data$tumor_stage)
# imputed_data$tumor_stage <- as.factor(imputed_data$tumor_stage)
# 
# # Create a combined dataset for plotting
# plot_data <- bind_rows(
#   data %>% select(tumor_stage) %>% mutate(Source = 'Original'),
#   imputed_data %>% select(tumor_stage) %>% mutate(Source = 'Imputed')
# )
# 
# # Plot both distributions
# ggplot(data, aes(x = tumor_stage)) +
#   geom_bar(position = "dodge") +
#   labs(title = "Distribution of Tumor Stage: Original vs Imputed",
#        x = "Tumor Stage",
#        y = "Count") +
#   theme_minimal()
```

# 1. Filtering
```{r}
# Check for outliers
imputed_data <- read_csv("imputed_data.csv")

library(isotree)

x<-isolation.forest(imputed_data, ndim = 1, sample_size = 256, max_depth = 8, ntrees = 100, missing_action = "fail", categ_split_type="single_categ", scoring_metric="density")

data_filtered <- imputed_data
data_filtered$pred <- predict(x, imputed_data)

# Create a data frame for plotting
scores_df <- data.frame(
  OutlierScore = data_filtered$pred
)

# Plot the distribution of outlier scores
ggplot(scores_df, aes(x = OutlierScore)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(x = "Outlier Score", y = "Frequency", title = "Distribution of Outlier Scores") +
  theme_minimal()

# Calculate mean and standard deviation
mean_score <- mean(data_filtered$pred, na.rm = TRUE)
sd_score <- sd(data_filtered$pred, na.rm = TRUE)

# Store the original number of rows before filtering
original_count <- nrow(data_filtered)

lower_threshold <- mean_score - 2*sd_score

# Filter out rows where the score is outside of 2 standard deviations from the mean
data_filtered <- data_filtered %>%
  filter(pred >= lower_threshold) %>%
  select(-pred)  # Remove the 'pred' column

# Store the number of rows after filtering
filtered_count <- nrow(data_filtered)

# Calculate the number of eliminated rows
eliminated_count <- original_count - filtered_count

# Print the number of eliminated variables
cat("Number of eliminated variables:", eliminated_count, "\n")
```

# 2. The random forest
```{r data tranformation for random forest, echo=FALSE}

# Load required packages
library(randomForest)
library(caTools)
library(caret)
library(nestedcv)
library(mlbench)
library(pbapply)
library(ROCR)
library(PRROC)
library(pROC)
library(gridExtra)

# Create a function to create the prognosis variable
categorise_prognosis <- function(months) {
  case_when(
    months <= 60 + (60 * 0.25) ~ "will_die_within_5_years",
    months > 60 + (60 * 0.25) ~ "will_die_within_10_years",
  )
}

# Process the data
categorical_data <- data_filtered %>%
  select(age_at_diagnosis:death_from_cancer) %>%
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

source("~/University/MSB/MSB1015/funPCA.R") # Change the directory as needed

# Call the function and plot
pca_plots_categorical <- pca_proximity_plot(categorical_data, "prognosis")
grid.arrange(grobs = pca_plots_categorical, ncol = 2)

genetic_data <- data_filtered %>%
  select(brca1:stmn2_mut, overall_survival_months, death_from_cancer, overall_survival) %>%
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function and plot
pca_plots_genetic <- pca_proximity_plot(genetic_data, "prognosis")
grid.arrange(grobs = pca_plots_genetic, ncol = 2)

complete_data <- data_filtered %>%
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function and plot
pca_plots_complete <- pca_proximity_plot(complete_data, "prognosis")
grid.arrange(grobs = pca_plots_complete, ncol = 2)
```

# 2.1 The random forest
```{r random forest, echo=FALSE}

# Set the seed for reproducibility of random results using the L'Ecuyer-CMRG method
set.seed(111, "L'Ecuyer-CMRG")

# Create a partition to split the data into training and testing sets
# The target variable for stratification is 'prognosis', with 70% of the data for training
trainIndex <- createDataPartition(categorical_data$prognosis, p = .7, 
                                  list = FALSE,  # Return indices as a matrix
                                  times = 1)     # Perform the partitioning once

# Subset the original data to create the training set using the indices from trainIndex
train <- categorical_data[trainIndex,]

# Subset the original data to create the testing set using the remaining indices
test  <- categorical_data[-trainIndex,]

# Ensure "prognosis" is a factor
train$prognosis <- as.factor(train$prognosis)
test$prognosis <- as.factor(test$prognosis)

# Convert train and test features to data frames to avoid tibble row name issues
train_features <- as.data.frame(train %>% select(-prognosis))
test_features <- as.data.frame(test %>% select(-prognosis))
y_train <- train$prognosis

train_sample <- randomsample(y_train, train_features)
y_train <- train_sample$y
x_train <- train_sample$x

# Define the resampling strategy for outer cross-validation
inner_control <- trainControl(method = "cv", number = 5, search = "grid", summaryFunction = twoClassSummary, classProbs = TRUE)

tuneGrid <- expand.grid(mtry = 2:8)

# The model with the highest AUC that I found was with 2000 trees and node size of 1. 
# When creating the model these values can be changed.
ntrees <- c(2000) 
nodesize <- c(1)

# List for the combinations: should have all the possible combinations between ntrees and nodesize.
list <- 1

# Define the grid for hyperparameter tuning
params <- expand.grid(ntrees = ntrees,
                      nodesize = nodesize)

store_models <- vector("list", nrow(params))

# Perform nested cross-validation with Random Forest
for(i in 1:nrow(params)) {
  nodesize <- params[i,2]
  ntree <- params[i,1]

  # Train the model with the specified parameters
  nested_result <- nestcv.train(
    x = x_train,
    y = y_train,
    method = "rf",
    savePredictions = "final",
    trControl = inner_control,  # Use the inner cross-validation control
    tuneGrid = tuneGrid,
    ntree = ntree,
    nodesize = nodesize,
    importance = TRUE,
    cv.cores = 8,
    metric = "ROC",
    filterFUN = rf_filter
  )

  # Assign the nested result to store_models and name it based on parameters
  store_models[[i]] <- nested_result
  names(store_models)[i] <- paste0("ntree_", ntree, "_nodesize_", nodesize)
}

```

# 2.2 Finding the best model
```{r best model, echo=FALSE}
# Initialize an empty list to store the metrics
metrics_list <- list()

# Loop over the indices in the provided list
for (i in seq_along(store_models)) {
  # Apply the metrics function to each model in store_models
  metrics_list[[names(store_models)[i]]] <- metrics(store_models[[i]])
}

# Combine the list of metrics into a dataframe
metrics_df <- do.call(rbind, metrics_list)
metrics_df <- as.data.frame(metrics_df)

# View the resulting dataframe
print(metrics_df)

# Find the model with the highest AUC
best_model_index <- which.max(metrics_df$AUC)

# Access the best model based on the highest ROC
best_model <- store_models[[best_model_index]]

# Get an overview of the best model
summary(best_model)
```

# 2.3 ROC and PR curves on the best model
```{r accuracy and plotting, echo=FALSE}
# Calculate accuracy on the train set for the best model
pred_table_training <- table(best_model$output$predy, best_model$output$testy)
train_confusion <- confusionMatrix(pred_table_training)
train_confusion

# Call the plotting function for the confusion matrix
source("~/University/MSB/MSB1015/confusionPlot.R") # Change the directory as needed
confusionPlot(train_confusion, title = "Confusion Matrix (Training Set)")

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_model, newdata = train, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"]

# Get the precision recall values
best_model$prc <- prc(best_model)

# Save precision and recall plots as PNG
png("precision_recall_plots_training.png", width = 1600, height = 800)

# Plot precision and recall
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1)
plot(best_model$roc, col = "red", main = "ROC Training", las = 1)
legend("bottomright", legend = paste0("AUC = ", signif(pROC::auc(best_model$roc), 3)), bty = 'n')
plot(best_model$prc, col = "red", main = "Precision-recall Training")
legend("topright", legend = paste0("AUC = ", round(best_model$prc$auc, 3)), bty = 'n')

# Reset plotting parameters
par(op)

# Close the PNG device
dev.off()

# Extract the importance of the variables for the best model
importance <- as.data.frame(best_model$final_fit$finalModel$importance) %>%
  mutate(will_die_within_10_years = will_die_within_10_years * 10,
         will_die_within_5_years = will_die_within_5_years * 10,
         column = row.names(best_model$final_fit$finalModel$importance))

importance_sorted <- importance[order(importance$will_die_within_10_years, decreasing = TRUE), ]
head(importance_sorted)

# Plot the importance of the variables
ggplot(importance, aes(x = reorder(column, will_die_within_10_years))) +
  geom_bar(aes(y = will_die_within_10_years, fill = "Will die within 10 years"), 
           position = "dodge", stat = "identity") +
  geom_bar(aes(y = will_die_within_5_years, fill = "Will die within 5 years"), 
           position = "dodge", stat = "identity") +
  scale_fill_manual(values = c("Will die within 10 years" = "skyblue1", 
                               "Will die within 5 years" = "tomato")) +
  coord_flip() +
  labs(x = "Feature", y = "Importance Score", title = "Variable Importance", fill = "Prognosis") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

# Extract the error rates (OOB)
err_rate <- as.data.frame(best_model$final_fit$finalModel$err.rate) %>%
  mutate(ntrees = 1:nrow(best_model$final_fit$finalModel$err.rate))

# Plot OOB
ggplot(err_rate, aes(x = ntrees, y = OOB)) +
  geom_line(color = "skyblue1", linewidth = 1) +
  labs(x = "Number of Trees", y = "OBB Error", title = "Out Of Bags (OOB) error") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

# 2.4 Testing
```{r testing, echo=FALSE}

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_model, newdata = test, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"]

# Generate the ROC curve
roc_curve <- roc(test$prognosis, positive_class_probs, levels = levels(test$prognosis), direction = ">")

# Create Precision-Recall curve
pr_curve <- pr.curve(scores.class0 = positive_class_probs, 
                     weights.class0 = as.numeric(test$prognosis == "will_die_within_10_years"), 
                     curve = TRUE)

# Save precision and recall plots as PNG
png("precision_recall_plots_train_test.png", width = 1600, height = 800)

# Plot combined graphs
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1)
plot(best_model$roc, col = "red", main = "ROC", las = 1)
plot(roc_curve, col = "blue", add = TRUE, main = "ROC", las = 1)
legend("bottomright", 
       legend = c(paste0("AUC (Training) = ", signif(pROC::auc(best_model$roc), 2)), 
                  paste0("AUC (Testing) = ", signif(roc_curve$auc, 2))), 
       col = c("red", "blue"), lty = 1, bty = 'n')

# Plot the Precision-Recall curve for the training data
plot(best_model$prc, col = "red", main = "Precision-Recall Curve", xlab = "Recall", ylab = "Precision")
lines(pr_curve$curve[, 3], pr_curve$curve[, 1], col = "blue")
legend("topright", 
       legend = c(paste0("AUC (Training) = ", round(best_model$prc$auc, 2)), 
                  paste0("AUC (Testing) = ", round(pr_curve$auc.integral, 2))), 
       col = c("red", "blue"), 
       lty = 1, 
       bty = 'n')

# Reset plotting parameters and close PNG
par(op)
dev.off()

# Find the best threshold from the ROC curve (maximizing sensitivity-specificity balance)
best_threshold <- as.numeric(coords(roc_curve, "best", ret = "threshold"))
print(paste("Best threshold based on ROC curve:", best_threshold))

# Apply the best threshold for classification
pred_best_threshold <- ifelse(positive_class_probs > best_threshold,
                              "will_die_within_10_years",
                              "will_die_within_5_years")

# Ensure the predicted classes are factors with the correct levels
pred_best_threshold <- factor(pred_best_threshold, levels = levels(test$prognosis))

# Calculate accuracy on the test set with the best threshold
pred_table <- table(pred_best_threshold, test$prognosis)
dimnames(pred_table) <- list(c("will_die_within_10_years", "will_die_within_5_years"), 
                             c("will_die_within_10_years", "will_die_within_5_years"))
test_confusion <- confusionMatrix(pred_table)
test_confusion

confusionPlot(test_confusion, title = "Confusion Matrix (Testing Set)")
```