### 2024-10-14 Code/MSB1015-Project.Rmd
Update MSB1015-Project.Rmd
Changes in the styling of the code.
Addition of for loops for the random forest process.
Minor updates to the figures.
Made the code more streamlined for multiple datasets.
Removed unused lines.

---
title: "MSB1015 Project"
author: "Carlo Alberto Zani"
date: "2024-10-14"
output: html_document
---

# 0. Setup
```{r setup, include=FALSE}
library(rstudioapi)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(GGally)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()

data <- read_csv("METABRIC_RNA_Mutation.csv")
```

# 0.1 Data Cleaning
```{r cleaning, echo=FALSE}
# Check for duplicates
any(duplicated(data))

# Check for missing values
missing_counts <- data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "missing_count") %>%
  filter(missing_count > 10)

# Plot where those missing values were found
ggplot(missing_counts, aes(x = reorder(column, missing_count), y = missing_count)) +
  geom_bar(stat = "identity", fill = "skyblue1") +
  coord_flip() +
  labs(x = "Column", y = "Missing Count", title = "Columns with More Than 10 Missing Values")

data$tumor_stage <- as.factor(data$tumor_stage)

data$neoplasm_histologic_grade <- as.factor(data$neoplasm_histologic_grade)

# Apply the modifications
data_filled <- data %>%
  # 1. Replace missing values with "Unknown" for all categorical columns
  mutate(
    oncotree_code = replace_na(oncotree_code, "Unknown"),
    cancer_type_detailed = replace_na(cancer_type_detailed, "Unknown")
  ) %>%
  # 3. Replace missing values with the median for "tumor_size"
  mutate(tumor_size = ifelse(is.na(tumor_size), median(tumor_size, na.rm = TRUE), tumor_size)) %>%
  # 4. Replace missing values with 0 for columns ending with "_mut"
  mutate(across(ends_with("_mut"), ~ replace_na(., 0))) %>%
  # 5. Convert the columns ending with "_mut" to a logical vector
  mutate(across(ends_with("_mut"), ~ ifelse(. != "0", 1, 0))) %>%
  # 6.
  rowwise() %>%
  mutate(mutation_count = sum(across(ends_with("_mut")) == 1))

# Check for missing values after cleaning
sum(is.na(data_filled))

# Select only the transcription data
pca_gx <- data_filled %>%
  select(brca1:ugt2b7)

# Perform PCA to check the columns with the highest amount of missing values
pca <- prcomp(pca_gx)

# Convert PCA results to data frame
pca_data <- as.data.frame(pca$x[, 1:3])
pca_data$tumor_stage <- as.factor(data_filled$tumor_stage)

# Create pairwise plots using ggpairs
ggpairs(pca_data,
  mapping = aes(color = tumor_stage),
  upper = list(continuous = wrap("points")),
  lower = list(continuous = wrap("points")),
  diag = list(continuous = wrap("barDiag"))
) +
  theme_minimal()
```

# 0.2 Imputation (this step will take some time)
```{r impute, echo=FALSE}
# Load necessary libraries
library(mice)

# Get rid of unwanted characters from the column names
colnames(data_filled) <- make.names(colnames(data_filled), unique = TRUE)

# Convert all character variables to factors
data_filled <- data_filled %>%
  mutate(across(where(is.character), as.factor))

# Identify categorical variables
fac_vars <- names(data_filled)[sapply(data_filled, is.factor)]

# Specify methods for all variables, setting CART for categorical variables
meth <- make.method(data_filled)
meth[fac_vars] <- "cart" # Use CART for all categorical variables

set.seed(111)

# Perform the imputation using the 'mice' package
imputation_result <- mice(data_filled,
  method = "cart", # 'meth' contains the method for categorical imputation (CART)
  m = 5 # 5 is the commonly chosen value for the number of multiple imputations to perform
)

imputed_data <- complete(imputation_result)

write.csv(imputed_data, file = "imputed_data.csv", row.names = FALSE)
```

# 0.3 Plots for imputation
```{r plots after imputation, echo=FALSE}
# Load the imputed data from the previous steps
imputed_data <- read_csv("imputed_data.csv")

# Convert 'tumor_stage' to a factor in both datasets
data$tumor_stage <- as.factor(data$tumor_stage)
imputed_data$tumor_stage <- as.factor(imputed_data$tumor_stage)

# Create a combined dataset for plotting
plot_data <- bind_rows(
  data %>% select(tumor_stage) %>% mutate(Source = "Original"),
  imputed_data %>% select(tumor_stage) %>% mutate(Source = "Imputed")
)

# Plot the imputed vs original
ggplot(plot_data, aes(x = tumor_stage, fill = Source)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of missing values for Tumor Stage: Original vs Imputed",
    x = "Tumor Stage",
    y = "Count"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Original" = "royalblue", "Imputed" = "firebrick1"))
```

# 1. Filtering
```{r isotree filtering, echo=FALSE}
# You might get an error when loading this package but just run it again and it should work
library(isotree)

# In the 'isolation.forest' function parameters were kept as recommended (not default).
isolation_result <- isolation.forest(imputed_data,
  ndim = 1, # It was recommended for data with many categorical variables like mine
  sample_size = 256, # This was the recommended value in most sources
  max_depth = 8, # ceiling(log2(256)) is the recommended value
  ntrees = 1000, # The recommended is 100 but the more trees the better usually.
  missing_action = "fail", # The imputed data has no missing values but in case it does the function would stop
  categ_split_type = "single_categ", # It was recommended for data with many categorical variables like mine
  scoring_metric = "depth" # The depth score gives the best performance when looking at literature
)

data_filtered <- imputed_data
data_filtered$pred <- predict(isolation_result, imputed_data)

# Create a data frame for plotting
scores_df <- data.frame(
  OutlierScore = data_filtered$pred
)

# Plot the distribution of outlier scores
ggplot(scores_df, aes(x = OutlierScore)) +
  geom_histogram(bins = 30, fill = "royalblue", color = "black") +
  labs(x = "Outlier Score", y = "Frequency", title = "Distribution of Outlier Scores") +
  theme_minimal()

# Store the original number of rows before filtering
original_count <- nrow(data_filtered)

# Filter out rows where the score is above 0.5 which is the standard value for outlier detection
data_filtered <- data_filtered %>%
  filter(pred <= 0.5) %>%
  select(-pred)

# Store the number of rows after filtering
filtered_count <- nrow(data_filtered)

# Calculate the number of eliminated rows
eliminated_count <- original_count - filtered_count

# Print the number of eliminated variables
cat("Number of eliminated variables:", eliminated_count, "\n")
```

# 2. The random forest
```{r data tranformation for random forest, echo=FALSE}
# Load required packages
library(randomForest)
library(caTools)
library(caret)
library(nestedcv)
library(mlbench)
library(pbapply)
library(ROCR)
library(PRROC)
library(pROC)
library(gridExtra)

# Create a function to form the prognosis variable
categorise_prognosis <- function(months) {
  case_when(
    months <= 60 + (60 * 0.25) ~ "will_die_within_5_years",
    months > 60 + (60 * 0.25) ~ "will_die_within_10_years",
  )
}

# Create the categorical dataframe
categorical_data <- data_filtered %>%
  select(age_at_diagnosis:death_from_cancer) %>% # Extract the categorical variables
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months)) # Call the function to create 'prognosis'
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

source("~/University/MSB/MSB1015/funPCA.R") # Change the directory as needed

# Call the function for the unsupervised random forest PCA
pca_plots_categorical <- pca_proximity_plot(categorical_data, "prognosis")
grid.arrange(grobs = pca_plots_categorical, ncol = 2)

# Create the genetic dataframe
genetic_data <- data_filtered %>%
  select(brca1:stmn2_mut, overall_survival_months, death_from_cancer, overall_survival) %>% # Extract the genetic variables and those needed to calculate prognosis
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function for the unsupervised random forest PCA (will take a few minutes)
pca_plots_genetic <- pca_proximity_plot(genetic_data, "prognosis")
grid.arrange(grobs = pca_plots_genetic, ncol = 2)

# Create the complete dataframe
complete_data <- data_filtered %>%
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function for the unsupervised random forest PCA (will take a few minutes)
pca_plots_complete <- pca_proximity_plot(complete_data, "prognosis")
grid.arrange(grobs = pca_plots_complete, ncol = 2)
```

# 2.1 The random forest function
```{r random forest function, echo=FALSE}
random_forest <- function(data, ntree, nrun) {
  # Set the seed for reproducibility using the L'Ecuyer-CMRG method
  set.seed(111, "L'Ecuyer-CMRG")

  # Initialize an empty list to store models from repeated runs
  store_models_repeated <- vector("list", nrun)

  # Loop to perform nested cross-validation nrun times
  for (repeat_run in 1:nrun) {
    # Create a partition to split the data into training and testing sets
    # The target variable for stratification is 'prognosis', with 70% of the data for training
    trainIndex <- createDataPartition(data$prognosis,
      p = .7,
      list = FALSE, # Return indices as a matrix
      times = 1
    ) # Perform the partitioning once

    # Subset the original data to create training and testing sets
    train <- data[trainIndex, ]
    test <- data[-trainIndex, ]

    # Ensure "prognosis" is a factor
    train$prognosis <- as.factor(train$prognosis)
    test$prognosis <- as.factor(test$prognosis)

    # Convert train features to data frames
    train_features <- as.data.frame(train %>% select(-prognosis))
    y_train <- train$prognosis

    # Random oversampling of the minority class from 'prognosis'
    train_sample <- randomsample(y_train, train_features)
    y_train <- train_sample$y
    x_train <- train_sample$x

    # Define the resampling strategy for outer cross-validation.
    inner_control <- trainControl(
      method = "cv", # Method was kept as cross validation as standard since it performed the best
      number = 5, # 5 folds were used for the inner folds as it gave the best result
      search = "grid", # A grid search was used to test the different tuning variables defined later
      summaryFunction = twoClassSummary, # Used to retrieve the values need for the ROC and PR curves
      classProbs = TRUE # Used to retrieve the values need for the ROC and PR curves
    ) # Everything else was kept standard and NULL as it was not needed.

    # Create a tune grid for the 'mtry' variable
    tuneGrid <- expand.grid(mtry = 2:8)

    # The model with the highest AUC that I found was with 2000 trees and node size of 1.
    nodesize <- c(1)
    params <- expand.grid(ntrees = ntree, nodesize = nodesize)

    # Store models for the current repeat
    store_models <- vector("list", nrow(params))

    # Nested cross-validation loop with Random Forest
    for (i in 1:nrow(params)) {
      nodesize <- params[i, 2]
      ntree <- params[i, 1]

      # Train the model with the specified parameters.
      nested_result <- nestcv.train(
        x = x_train,
        y = y_train,
        method = "rf", # Random forest from the 'randomForest' function
        savePredictions = "final", # 'savePredictions' was kept as final so that it would save the final model with the best scores.
        trControl = inner_control, # Use the inner cross-validation control
        tuneGrid = tuneGrid, # The tune grid is used for the tuning
        ntree = ntree, # One of the tuned variables
        nodesize = nodesize, # One of the tuned variables
        importance = TRUE, # The importance of the variables is scored for later visualisation
        cv.cores = 8, # Here you can set how many cores to use for parallelisation
        metric = "ROC", # The 'ROC' metric is used to optimise the models as it is a better representation of performance compared to accuracy.
        filterFUN = rf_filter # Filtering of the variables helps, especially with the genetic and full databases
      )

      # Store the result
      store_models[[i]] <- nested_result
      names(store_models)[i] <- paste0("ntree_", ntree, "_nodesize_", nodesize)
    }

    # Assign the models from this run to store_models_repeated
    store_models_repeated[[repeat_run]] <- store_models
  }

  # Initialize an empty list to store the metrics for all repetitions
  metrics_list_repeated <- list()

  # Loop over all repetitions and calculate metrics for each model
  for (run in seq_along(store_models_repeated)) {
    store_models <- store_models_repeated[[run]]

    # Initialize an empty list to store the metrics for this repetition
    metrics_list <- list()

    # Loop over the models in the current repetition
    for (i in seq_along(store_models)) {
      # Apply the metrics function to each model
      metrics_list[[names(store_models)[i]]] <- metrics(store_models[[i]])
    }

    # Combine the list of metrics into a dataframe
    metrics_df <- do.call(rbind, metrics_list)
    metrics_df <- as.data.frame(metrics_df)

    # Store the metrics for this repetition
    metrics_list_repeated[[run]] <- metrics_df
  }

  # Find the best model across all repetitions
  best_model_overall <- NULL
  best_auc <- 0

  for (run in seq_along(metrics_list_repeated)) {
    metrics_df <- metrics_list_repeated[[run]]

    # Find the model with the highest AUC in this repetition
    best_model_index <- which.max(metrics_df$AUC)

    # Check if it's the best model overall
    if (metrics_df$AUC[best_model_index] > best_auc) {
      best_auc <- metrics_df$AUC[best_model_index]
      best_model_overall <- store_models_repeated[[run]][[best_model_index]]
    }
  }

  # Get an overview of the best model
  print("Best Model Summary")
  summary(best_model_overall)

  # Return the best model and the list of models for each repetition
  return(list(best_model_overall = best_model_overall, store_models_repeated = store_models_repeated, train = train, test = test))
}
```

# 2.2 Calling the random forest function
```{r perform random forest, echo=FALSE}
# Categorical results
results_categorical <- random_forest(categorical_data, ntree = 10, nrun = 2)

# Genetic results
results_genetic <- random_forest(genetic_data, ntree = 10, nrun = 2)

# Complete results
results_complete <- random_forest(complete_data, ntree = 10, nrun = 2)
```

# 3.1 Create a function to the scores for all the models
```{r model performance function, echo=FALSE}
# Create the function to extract the metrics from the result list
get_final_metrics <- function(results_data) {
  # Initialize an empty list to store all metrics from all repetitions
  all_metrics_list <- list()

  # Loop over all repetitions and calculate metrics for each model
  for (run in seq_along(results_data[[2]])) {
    store_models <- results_data[[2]][[run]]

    # Initialize an empty list to store the metrics for this repetition
    metrics_list <- list()

    # Loop over the models in the current repetition
    for (i in seq_along(store_models)) {
      # Apply the metrics function to each model
      metrics_list[[names(store_models)[i]]] <- metrics(store_models[[i]])
    }

    # Combine the list of metrics into a dataframe
    metrics_df <- do.call(rbind, metrics_list)
    metrics_df <- as.data.frame(metrics_df)

    # Add the run information to the metrics dataframe
    metrics_df$Run <- run # Adding a column to indicate which repetition the metrics belong to

    # Store the metrics for this repetition
    all_metrics_list[[run]] <- metrics_df
  }

  # Combine all metrics from all runs into a single dataframe
  final_metrics_df <- do.call(rbind, all_metrics_list)

  # Transpose the dataframe
  t_final_metrics_df <- t(final_metrics_df)

  # Convert the matrix to a data frame and keep only the 1st and 3rd rows (AUC, Balanced Accuracy)
  t_final_metrics_df <- as.data.frame(t_final_metrics_df[c(1, 3), ])

  # Compute the average across the rows
  t_final_metrics_df <- t_final_metrics_df %>%
    rowwise() %>%
    mutate(average = mean(c_across(everything()), na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(Metric = c("AUC", "Balanced Accuracy"))

  # Return the final metrics dataframe
  return(t_final_metrics_df)
}
```

# 3.2 Call the function to find the best model
```{r create the metric dataframes, echo=FALSE}
# Categorical metrics
metrics_categorical <- get_final_metrics(results_categorical)
print(metrics_categorical)

# Genetic metrics
metrics_genetic <- get_final_metrics(results_genetic)
print(metrics_genetic)

# Complete metrics
metrics_complete <- get_final_metrics(results_complete)
print(metrics_complete)

# Add class column to each dataframe
metrics_categorical$class <- "categorical"
metrics_genetic$class <- "genetic"
metrics_complete$class <- "complete"

# Combine all dataframes into one
all_metrics <- bind_rows(metrics_categorical, metrics_genetic, metrics_complete)

# Extract only the rows corresponding to AUC
auc_metrics <- all_metrics %>%
  filter(Metric == "AUC") %>%
  filter(average == max(average))

# Extract and print the class of the best model
best_model_class <- auc_metrics$class

# Create the name of the dataset corresponding to the best model class
best_class_name <- paste("results_", best_model_class, sep = "")

# Print the best class name
print(best_class_name)

# Get the the name of the data with the best model
best_class_data <- get(best_class_name)
```

# 4. ROC and PR curves on the best model from the best class
```{r accuracy and plotting, echo=FALSE}
# Calculate accuracy on the train set for the best model
pred_table_training <- table(best_class_data$best_model_overall$output$predy, best_class_data$best_model_overall$output$testy)
train_confusion <- confusionMatrix(pred_table_training)
train_confusion

# Call the plotting function for the confusion matrix
source("~/University/MSB/MSB1015/confusionPlot.R") # Change the directory as needed
confusionPlot(train_confusion, title = "Confusion Matrix (Training Set)")

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = best_class_data$train, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"]

# Get the precision recall values
best_class_data$best_model_overall$prc <- prc(best_class_data$best_model_overall)

# Save precision and recall plots as PNG
png("precision_recall_plots_training.png", width = 1600, height = 800)

# Plot precision and recall
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1)
plot(best_class_data$best_model_overall$roc, col = "red", main = "ROC Training", las = 1)
legend("bottomright", legend = paste0("AUC = ", signif(pROC::auc(best_class_data$best_model_overall$roc), 3)), bty = "n")
plot(best_class_data$best_model_overall$prc, col = "red", main = "Precision-recall Training")
legend("topright", legend = paste0("AUC = ", round(best_class_data$best_model_overall$prc$auc, 3)), bty = "n")

# Reset plotting parameters
par(op)

# Close the PNG device
dev.off()

# Extract the importance of the variables for the best model
importance <- as.data.frame(best_class_data$best_model_overall$final_fit$finalModel$importance) %>%
  mutate(
    will_die_within_10_years = will_die_within_10_years * 10,
    will_die_within_5_years = will_die_within_5_years * 10,
    column = row.names(best_class_data$best_model_overall$final_fit$finalModel$importance)
  )

importance_sorted <- importance[order(importance$will_die_within_10_years, decreasing = TRUE), ]
head(importance_sorted)

# Plot the importance of the variables
ggplot(importance, aes(x = reorder(column, will_die_within_10_years))) +
  geom_bar(aes(y = will_die_within_10_years, fill = "Will die within 10 years"),
    position = "dodge", stat = "identity"
  ) +
  geom_bar(aes(y = will_die_within_5_years, fill = "Will die within 5 years"),
    position = "dodge", stat = "identity"
  ) +
  scale_fill_manual(values = c(
    "Will die within 10 years" = "skyblue1",
    "Will die within 5 years" = "tomato"
  )) +
  coord_flip() +
  labs(x = "Feature", y = "Importance Score", title = "Variable Importance", fill = "Prognosis") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

# Extract the error rates (OOB)
err_rate <- as.data.frame(best_class_data$best_model_overall$final_fit$finalModel$err.rate) %>%
  mutate(ntrees = 1:nrow(best_class_data$best_model_overall$final_fit$finalModel$err.rate))

# Plot OOB
ggplot(err_rate, aes(x = ntrees, y = OOB)) +
  geom_line(color = "skyblue1", linewidth = 1) +
  labs(x = "Number of Trees", y = "OBB Error", title = "Out Of Bags (OOB) error") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

# 5. Testing on the best model
```{r testing, echo=FALSE}
# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = best_class_data$test, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"]

# Generate the ROC curve
roc_curve <- roc(best_class_data$test$prognosis, positive_class_probs, levels = levels(best_class_data$test$prognosis), direction = ">")

# Create Precision-Recall curve
pr_curve <- pr.curve(
  scores.class0 = positive_class_probs,
  weights.class0 = as.numeric(best_class_data$test$prognosis == "will_die_within_10_years"),
  curve = TRUE
)

# Save precision and recall plots as PNG
png("precision_recall_plots_train_test.png", width = 1600, height = 800)

# Plot combined graphs
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1)
plot(best_class_data$best_model_overall$roc, col = "red", main = "ROC", las = 1)
plot(roc_curve, col = "blue", add = TRUE, main = "ROC", las = 1)
legend("bottomright",
  legend = c(
    paste0("AUC (Training) = ", signif(pROC::auc(best_class_data$best_model_overall$roc), 2)),
    paste0("AUC (Testing) = ", signif(roc_curve$auc, 2))
  ),
  col = c("red", "blue"), lty = 1, bty = "n"
)

# Plot the Precision-Recall curve for the training data
plot(best_class_data$best_model_overall$prc, col = "red", main = "Precision-Recall Curve", xlab = "Recall", ylab = "Precision")
lines(pr_curve$curve[, 3], pr_curve$curve[, 1], col = "blue")
legend("topright",
  legend = c(
    paste0("AUC (Training) = ", round(best_class_data$best_model_overall$prc$auc, 2)),
    paste0("AUC (Testing) = ", round(pr_curve$auc.integral, 2))
  ),
  col = c("red", "blue"),
  lty = 1,
  bty = "n"
)

# Reset plotting parameters and close PNG
par(op)
dev.off()

# Find the best threshold from the ROC curve (maximizing sensitivity-specificity balance)
best_threshold <- as.numeric(coords(roc_curve, "best", ret = "threshold"))
print(paste("Best threshold based on ROC curve:", best_threshold))

# Apply the best threshold for classification
pred_best_threshold <- ifelse(positive_class_probs > best_threshold,
  "will_die_within_10_years",
  "will_die_within_5_years"
)

# Ensure the predicted classes are factors with the correct levels
pred_best_threshold <- factor(pred_best_threshold, levels = levels(best_class_data$test$prognosis))

# Calculate accuracy on the test set with the best threshold
pred_table <- table(pred_best_threshold, best_class_data$test$prognosis)
dimnames(pred_table) <- list(
  c("will_die_within_10_years", "will_die_within_5_years"),
  c("will_die_within_10_years", "will_die_within_5_years")
)
test_confusion <- confusionMatrix(pred_table)
test_confusion

confusionPlot(test_confusion, title = "Confusion Matrix (Testing Set)")
```