### 2024-10-18 Code/MSB1015-Project.Rmd
Update MSB1015-Project.Rmd
Added a chunk to find the matching train and test splits with the best model

---
title: "MSB1015 Project"
author: "Carlo Alberto Zani"
date: "2024-10-14"
output: html_document
---

# 0. Setup
```{r setup, include=FALSE}
# List of required packages
required_packages <- c(
  "rstudioapi", "readr", "dplyr", "ggplot2", "tidyr",
  "stringr", "GGally", "mice", "isotree", "randomForest",
  "caTools", "caret", "nestedcv", "mlbench", "pbapply",
  "ROCR", "PRROC", "pROC", "gridExtra"
)

# Check and install missing packages
new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) {
  install.packages(new_packages)
}

# Load libraries
library(rstudioapi)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(GGally)
library(mice)
library(randomForest)
library(caTools)
library(caret)
library(nestedcv)
library(mlbench)
library(pbapply)
library(ROCR)
library(PRROC)
library(pROC)
library(gridExtra)

# Set the working directory to the one where this code is
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()

# Load the data
data <- read_csv("METABRIC_RNA_Mutation.csv")
```

# 0.1 Data Cleaning
```{r cleaning, echo=FALSE}
# Check for duplicates
any(duplicated(data))

# Check for missing values
missing_counts <- data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>% # Summarise missing values per column
  pivot_longer(everything(), names_to = "column", values_to = "missing_count") %>% # Reshape into long format
  filter(missing_count > 10) # Filter columns with more than 10 missing values

# Plot where those missing values were found
ggplot(missing_counts, aes(x = reorder(column, missing_count), y = missing_count)) +
  geom_bar(stat = "identity", fill = "royalblue") +
  coord_flip() +
  labs(x = "Column", y = "Missing Count", title = "Columns with More Than 10 Missing Values") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_line(color = "gray90"),
    axis.text.y = element_text(size = 14),  
    axis.text.x = element_text(size = 12),  
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 15, face = "bold")
  )

# Convert specific columns to factors
data$tumor_stage <- as.factor(data$tumor_stage)
data$neoplasm_histologic_grade <- as.factor(data$neoplasm_histologic_grade)

# Apply the modifications
data_filled <- data %>%
  # 1. Replace missing values with "Unknown" for specific categorical columns (not suitable for imputation)
  mutate(
    oncotree_code = replace_na(oncotree_code, "Unknown"),
    cancer_type_detailed = replace_na(cancer_type_detailed, "Unknown")
  ) %>%
  # 2. Replace missing values in "tumor_size" with the median value of the column
  mutate(tumor_size = ifelse(is.na(tumor_size), median(tumor_size, na.rm = TRUE), tumor_size)) %>%
  # 3. Replace missing values with 0 for all columns that end with "_mut"
  mutate(across(ends_with("_mut"), ~ replace_na(., 0))) %>%
  # 4. Convert values in columns ending with "_mut" to a binary vector (values not equal to "0" are converted to 1)
  mutate(across(ends_with("_mut"), ~ ifelse(. != "0", 1, 0))) %>%
  # 5. Calculate the "mutation_count" by summing the binary values (1s) in columns that end with "_mut" for each row
  rowwise() %>%
  mutate(mutation_count = sum(across(ends_with("_mut")) == 1))

# Check for missing values after cleaning
sum(is.na(data_filled))

# Select only the transcription data
pca_gx <- data_filled %>%
  select(brca1:ugt2b7)

# Perform PCA to check the columns with the highest amount of missing values
pca <- prcomp(pca_gx, center = TRUE, scale. = TRUE)

# Convert PCA results to data frame
pca_data <- as.data.frame(pca$x[, 1:3])
pca_data$tumor_stage <- as.factor(data_filled$tumor_stage)

# Create pairwise plots using ggpairs
ggpairs(pca_data,
  mapping = aes(color = tumor_stage),
  upper = list(continuous = wrap("points")),
  lower = list(continuous = wrap("points")),
  diag = list(continuous = wrap("barDiag"))
) +
  theme_minimal()
```

# 0.2 Imputation (this step will take some time)
```{r impute, echo=FALSE}
# Get rid of unwanted characters from the column names (the sign '+' from one of the variables)
colnames(data_filled) <- make.names(colnames(data_filled), unique = TRUE)

# Convert all character variables to factors
data_filled <- data_filled %>%
  mutate(across(where(is.character), as.factor))

set.seed(111)

# Perform the imputation using the 'mice' package
imputation_result <- mice(data_filled,
  method = "cart", # 'meth' contains the method for categorical imputation (CART)
  m = 5 # 5 is the commonly chosen value for the number of multiple imputations to perform
)

# Collect the results
imputed_data <- complete(imputation_result)

# Save the imputed data in the working directory
write.csv(imputed_data, file = "imputed_data.csv", row.names = FALSE)
```

# 0.3 Plots for imputation
```{r plots after imputation, echo=FALSE}
# Load the imputed data from the previous steps
imputed_data <- read_csv("imputed_data.csv")

# Convert 'tumor_stage' to a factor in both datasets
data$tumor_stage <- as.factor(data$tumor_stage)
imputed_data$tumor_stage <- as.factor(imputed_data$tumor_stage)

# Create a combined dataset for plotting
plot_data <- bind_rows(
  data %>% select(tumor_stage) %>% mutate(Source = "Original"),
  imputed_data %>% select(tumor_stage) %>% mutate(Source = "Imputed")
)

# Plot the imputed vs original
ggplot(plot_data, aes(x = tumor_stage, fill = Source)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of missing values for Tumor Stage: Original vs Imputed",
    x = "Tumor Stage",
    y = "Count"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Original" = "royalblue", "Imputed" = "firebrick1"))
```

# 1. Filtering
```{r isotree filtering, echo=FALSE}
# You might get an error when loading this package but just run it again and it should work
library(isotree)

# In the 'isolation.forest' function parameters were kept as recommended (not default).
isolation_result <- isolation.forest(imputed_data,
  ndim = 1, # It was recommended for data with many categorical variables like mine
  sample_size = 256, # This was the recommended value in most sources
  max_depth = 8, # ceiling(log2(256)) is the recommended value
  ntrees = 1000, # The recommended is 100 but the more trees the better usually.
  missing_action = "fail", # The imputed data has no missing values but in case it does the function would stop
  categ_split_type = "single_categ", # It was recommended for data with many categorical variables like mine
  scoring_metric = "depth" # The depth score gives the best performance when looking at literature
)

data_filtered <- imputed_data
data_filtered$pred <- predict(isolation_result, imputed_data)

# Create a data frame for plotting
scores_df <- data.frame(
  OutlierScore = data_filtered$pred
)

# Plot the distribution of outlier scores
ggplot(scores_df, aes(x = OutlierScore)) +
  geom_histogram(bins = 30, fill = "royalblue", color = "black") +
  labs(x = "Outlier Score", y = "Frequency", title = "Distribution of Outlier Scores") +
  theme_minimal()

# Store the original number of rows before filtering
original_count <- nrow(data_filtered)

# Filter out rows where the score is above 0.5 which is the standard value for outlier detection
data_filtered <- data_filtered %>%
  filter(pred <= 0.5) %>%
  select(-pred)

# Store the number of rows after filtering
filtered_count <- nrow(data_filtered)

# Calculate the number of eliminated rows
eliminated_count <- original_count - filtered_count

# Print the number of eliminated variables
cat("Number of eliminated variables:", eliminated_count, "\n")
```

# 2. The random forest
```{r data tranformation for random forest, echo=FALSE}
# Create a function to form the prognosis variable based on overall survival months
categorise_prognosis <- function(months) {
  case_when(
    months <= 60 + (60 * 0.25) ~ "will_die_within_5_years",
    months > 60 + (60 * 0.25) ~ "will_die_within_10_years"
  )
}

# Create the categorical dataframe
categorical_data <- data_filtered %>%
  select(age_at_diagnosis:death_from_cancer) %>% # Extract the categorical variables
  filter(death_from_cancer != "Died of Other Causes") %>% # Remove records for non-cancer deaths
  filter(overall_survival == 0) %>% # Keep only the rows where overall survival is 0
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months)) # Apply function to create 'prognosis'
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>% # Drop unnecessary columns
  as.data.frame() # Convert to a dataframe

# Load the function for PCA
source("~/University/MSB/MSB1015/funPCA.R") # Change the directory as needed

# Call the function for the unsupervised random forest PCA
pca_plots_categorical <- pca_proximity_plot(categorical_data, "prognosis")
grid.arrange(grobs = pca_plots_categorical, ncol = 2)

# Create the genetic dataframe
genetic_data <- data_filtered %>%
  select(brca1:stmn2_mut, overall_survival_months, death_from_cancer, overall_survival) %>% # Extract the genetic variables and those needed to calculate prognosis
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function for the unsupervised random forest PCA (will take a few minutes)
pca_plots_genetic <- pca_proximity_plot(genetic_data, "prognosis")
grid.arrange(grobs = pca_plots_genetic, ncol = 2)

# Create the complete dataframe
complete_data <- data_filtered %>%
  filter(death_from_cancer != "Died of Other Causes") %>%
  filter(overall_survival == 0) %>%
  mutate(
    prognosis = as.factor(categorise_prognosis(overall_survival_months))
  ) %>%
  select(-death_from_cancer, -overall_survival, -overall_survival_months) %>%
  as.data.frame()

# Call the function for the unsupervised random forest PCA (will take a few minutes)
pca_plots_complete <- pca_proximity_plot(complete_data, "prognosis")
grid.arrange(grobs = pca_plots_complete, ncol = 2)
```

# 2.1 The random forest function
```{r random forest function, echo=FALSE}
# Create the function to perform random forest classification
random_forest <- function(data, ntree, nrun) {
  # Set the seed for reproducibility using the L'Ecuyer-CMRG method
  set.seed(111, "L'Ecuyer-CMRG")

  # Initialize an empty list to store models from repeated runs
  store_models_repeated <- vector("list", nrun)

  # Initialize lists to store all train and test splits
  train_splits <- vector("list", nrun)
  test_splits <- vector("list", nrun)

  # Loop to perform nested cross-validation nrun times
  for (repeat_run in 1:nrun) {
    # Create a partition to split the data into training and testing sets
    # The target variable for stratification is 'prognosis', with 70% of the data for training
    trainIndex <- createDataPartition(data$prognosis,
      p = .7,
      list = FALSE, # Return indices as a matrix
      times = 1
    ) # Perform the partitioning once

    # Subset the original data to create training and testing sets
    train <- data[trainIndex, ]
    test <- data[-trainIndex, ]

    # Store the train and test sets for this run
    train_splits[[repeat_run]] <- train
    test_splits[[repeat_run]] <- test

    # Ensure "prognosis" is a factor
    train$prognosis <- as.factor(train$prognosis)
    test$prognosis <- as.factor(test$prognosis)

    # Convert train features to data frames
    train_features <- as.data.frame(train %>% select(-prognosis))
    y_train <- train$prognosis

    # Random oversampling of the minority class from 'prognosis'
    train_sample <- randomsample(y_train, train_features)
    y_train <- train_sample$y
    x_train <- train_sample$x

    # Define the resampling strategy for outer cross-validation.
    inner_control <- trainControl(
      method = "cv", # Method was kept as cross validation as standard since it performed the best
      number = 5, # 5 folds were used for the inner folds as it gave the best result
      search = "grid", # A grid search was used to test the different tuning variables defined later
      summaryFunction = twoClassSummary, # Used to retrieve the values need for the ROC and PR curves
      classProbs = TRUE # Used to retrieve the values need for the ROC and PR curves
    ) # Everything else was kept standard and NULL as it was not needed.

    # Create a tune grid for the 'mtry' variable
    tuneGrid <- expand.grid(mtry = 2:8)

    # The model with the highest AUC that I found was with 2000 trees and node size of 1.
    nodesize <- c(1)
    params <- expand.grid(ntrees = ntree, nodesize = nodesize)

    # Store models for the current repeat
    store_models <- vector("list", nrow(params))

    # Nested cross-validation loop with Random Forest
    for (i in 1:nrow(params)) {
      nodesize <- params[i, 2]
      ntree <- params[i, 1]

      # Train the model with the specified parameters.
      nested_result <- nestcv.train(
        x = x_train,
        y = y_train,
        method = "rf", # Random forest from the 'randomForest' function
        savePredictions = "final", # 'savePredictions' was kept as final so that it would save the final model with the best scores.
        trControl = inner_control, # Use the inner cross-validation control
        tuneGrid = tuneGrid, # The tune grid is used for the tuning
        ntree = ntree, # One of the tuned variables
        nodesize = nodesize, # One of the tuned variables
        importance = TRUE, # The importance of the variables is scored for later visualisation
        cv.cores = 8, # Here you can set how many cores to use for parallelisation
        metric = "ROC", # The 'ROC' metric is used to optimise the models as it is a better representation of performance compared to accuracy.
        filterFUN = rf_filter # Filtering of the variables helps, especially with the genetic and full databases
      )

      # Store the result
      store_models[[i]] <- nested_result
      names(store_models)[i] <- paste0("ntree_", ntree, "_nodesize_", nodesize)
    }

    # Assign the models from this run to store_models_repeated
    store_models_repeated[[repeat_run]] <- store_models
  }

  # Initialize an empty list to store the metrics for all repetitions
  metrics_list_repeated <- list()

  # Loop over all repetitions and calculate metrics for each model
  for (run in seq_along(store_models_repeated)) {
    store_models <- store_models_repeated[[run]]

    # Initialize an empty list to store the metrics for this repetition
    metrics_list <- list()

    # Loop over the models in the current repetition
    for (i in seq_along(store_models)) {
      # Apply the metrics function to each model
      metrics_list[[names(store_models)[i]]] <- metrics(store_models[[i]])
    }

    # Combine the list of metrics into a dataframe
    metrics_df <- do.call(rbind, metrics_list)
    metrics_df <- as.data.frame(metrics_df)

    # Store the metrics for this repetition
    metrics_list_repeated[[run]] <- metrics_df
  }

  # Find the best model across all repetitions
  best_model_overall <- NULL
  best_auc <- 0

  for (run in seq_along(metrics_list_repeated)) {
    metrics_df <- metrics_list_repeated[[run]]

    # Find the model with the highest AUC in this repetition
    best_model_index <- which.max(metrics_df$AUC)

    # Check if it's the best model overall
    if (metrics_df$AUC[best_model_index] > best_auc) {
      best_auc <- metrics_df$AUC[best_model_index]
      best_model_overall <- store_models_repeated[[run]][[best_model_index]]
    }
  }

  # Get an overview of the best model
  print("Best Model Summary")
  summary(best_model_overall)

  # Return the best model and the list of models for each repetition
  return(list(best_model_overall = best_model_overall, 
              store_models_repeated = store_models_repeated, 
              train_splits = train_splits, 
              test_splits = test_splits
              ))
}
```

# 2.2 Calling the random forest function
```{r perform random forest, echo=FALSE}
# Categorical results
results_categorical <- random_forest(categorical_data, ntree = 1000, nrun = 5)

# Genetic results
results_genetic <- random_forest(genetic_data, ntree = 10, nrun = 1)

# Complete results
results_complete <- random_forest(complete_data, ntree = 10, nrun = 1)
```

# 3.1 Create a function to the scores for all the models
```{r model performance function, echo=FALSE}
# Create the function to extract the metrics from the result list
get_final_metrics <- function(results_data) {
  # Initialize an empty list to store all metrics from all repetitions
  all_metrics_list <- list()

  # Loop over all repetitions and calculate metrics for each model
  for (run in seq_along(results_data[[2]])) {
    store_models <- results_data[[2]][[run]]

    # Initialize an empty list to store the metrics for this repetition
    metrics_list <- list()

    # Loop over the models in the current repetition
    for (i in seq_along(store_models)) {
      # Apply the metrics function to each model
      metrics_list[[names(store_models)[i]]] <- metrics(store_models[[i]])
    }

    # Combine the list of metrics into a dataframe
    metrics_df <- do.call(rbind, metrics_list)
    metrics_df <- as.data.frame(metrics_df)

    # Add the run information to the metrics dataframe
    metrics_df$Run <- run # Adding a column to indicate which repetition the metrics belong to

    # Store the metrics for this repetition
    all_metrics_list[[run]] <- metrics_df
  }

  # Combine all metrics from all runs into a single dataframe
  final_metrics_df <- do.call(rbind, all_metrics_list)

  # Transpose the dataframe
  t_final_metrics_df <- t(final_metrics_df)

  # Convert the matrix to a data frame and keep only the 1st and 3rd rows (AUC, Balanced Accuracy)
  t_final_metrics_df <- as.data.frame(t_final_metrics_df[c(1, 3), ])

  # Compute the average across the rows
  t_final_metrics_df <- t_final_metrics_df %>%
    rowwise() %>% # Perform row-wise operations
    mutate(average = mean(c_across(everything()), na.rm = TRUE)) %>% # Calculate row-wise average
    ungroup() %>%
    mutate(Metric = c("AUC", "Balanced Accuracy")) # Add a Metric column

  # Return the final metrics dataframe
  return(t_final_metrics_df)
}
```

# 3.2 Call the function to find the best model
```{r create the metric dataframes, echo=FALSE}
# Categorical metrics
metrics_categorical <- get_final_metrics(results_categorical)
print(metrics_categorical)

# Genetic metrics
metrics_genetic <- get_final_metrics(results_genetic)
print(metrics_genetic)

# Complete metrics
metrics_complete <- get_final_metrics(results_complete)
print(metrics_complete)

# Add class column to each dataframe
metrics_categorical$class <- "categorical"
metrics_genetic$class <- "genetic"
metrics_complete$class <- "complete"

# Combine all dataframes into one
all_metrics <- bind_rows(metrics_categorical, metrics_genetic, metrics_complete)

# Extract only the rows corresponding to AUC
auc_metrics <- all_metrics %>%
  filter(Metric == "AUC") %>%
  filter(average == max(average))

# Extract and print the class of the best model
best_model_class <- auc_metrics$class

# Create the name of the dataset corresponding to the best model class
best_class_name <- paste("results_", best_model_class, sep = "")

# Print the best class name
print(best_class_name)

# Get the the name of the data with the best model
best_class_data <- get(best_class_name)
```

# 3.3 Finding the matching splits for the best model
```{r finding matching splits}
# Get the AUC value of the best model
best_auc <- best_class_data$best_model_overall$roc$auc

# Initialize variables to store the matching run and model
matching_run <- NULL
matching_model <- NULL
matching_index <- NULL

# Loop over all runs in store_models_repeated
for (run_index in seq_along(best_class_data$store_models_repeated)) {
  # Get the current run's list of models
  current_run_models <- best_class_data$store_models_repeated[[run_index]]

  # Loop over all models in the current run and compare their AUC values
  for (model_index in seq_along(current_run_models)) {
    model_name <- names(current_run_models)[model_index]
    model_auc <- current_run_models[[model_name]]$roc$auc

    # Check if the AUC matches the best model's AUC
    if (model_auc == best_auc) {
      matching_run <- run_index
      matching_model <- model_name
      matching_index <- model_index
      break
    }
  }

  # Break out of the outer loop if a match is found
  if (!is.null(matching_run)) {
    break
  }
}

# If a match is found, print the run, model name, and retrieve the corresponding train/test splits
if (!is.null(matching_run) && !is.null(matching_model)) {
  print(paste("Matching model found in run:", matching_run, "model:", matching_model))

  # Retrieve the corresponding train and test splits
  matching_train_split <- best_class_data$train_splits[[matching_run]]
  matching_test_split <- best_class_data$test_splits[[matching_run]]

  # Print or use the splits as needed
  print("Found Matching Splits")

} else {
  print("No matching model found")
}

```

# 4. ROC and PR curves on the best model from the best class
```{r accuracy and plotting, echo=FALSE}
# Calculate accuracy on the train set for the best model
pred_table_training <- table(best_class_data$best_model_overall$output$predy, best_class_data$best_model_overall$output$testy) # Create prediction table
train_confusion <- confusionMatrix(pred_table_training) # Compute confusion matrix
train_confusion

# Call the plotting function for the confusion matrix
source("~/University/MSB/MSB1015/confusionPlot.R") # Load external plot function (update directory if needed)
confusionPlot(train_confusion, title = "Confusion Matrix (Training Set)") # Plot confusion matrix

# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = matching_train_split, type = "prob") # Predict probabilities
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Extract probabilities for "will_die_within_10_years"

# Get the precision-recall values
best_class_data$best_model_overall$prc <- prc(best_class_data$best_model_overall) # Calculate precision-recall

# Save precision and recall plots as PNG
png("precision_recall_plots_training.png", width = 1600, height = 800)

# Plot precision and recall
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1) # Set up the plotting layout
plot(best_class_data$best_model_overall$roc, col = "red", main = "ROC Training", las = 1) # Plot ROC curve
legend("bottomright", legend = paste0("AUC = ", signif(pROC::auc(best_class_data$best_model_overall$roc), 2)), bty = "n") # Add AUC legend
plot(best_class_data$best_model_overall$prc, col = "red", main = "Precision-recall Training") # Plot Precision-Recall curve
legend("topright", legend = paste0("AUC = ", round(best_class_data$best_model_overall$prc$auc, 2)), bty = "n") # Add AUC legend for PRC

# Reset plotting parameters
par(op)

# Close the PNG device
dev.off()

# Extract the importance of the variables for the best model
importance <- as.data.frame(best_class_data$best_model_overall$final_fit$finalModel$importance) %>% # Get variable importance
  mutate(
    will_die_within_10_years = will_die_within_10_years * 10, # Scale the importance scores
    will_die_within_5_years = will_die_within_5_years * 10,
    column = row.names(best_class_data$best_model_overall$final_fit$finalModel$importance) # Extract feature names
  )

importance_sorted <- importance[order(importance$will_die_within_10_years, decreasing = TRUE), ] # Sort by importance
head(importance_sorted)

# Plot the importance of the variables
ggplot(importance, aes(x = reorder(column, will_die_within_10_years))) +
  geom_bar(aes(y = will_die_within_10_years, fill = "Will die within 10 years"),
    position = "dodge", stat = "identity"
  ) +
  geom_bar(aes(y = will_die_within_5_years, fill = "Will die within 5 years"),
    position = "dodge", stat = "identity"
  ) +
  scale_fill_manual(values = c(
    "Will die within 10 years" = "skyblue1",
    "Will die within 5 years" = "tomato"
  )) +
  coord_flip() +
  labs(x = "Feature", y = "Importance Score", title = "Variable Importance", fill = "Prognosis") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) # Create variable importance plot

# Extract the error rates (OOB)
err_rate <- as.data.frame(best_class_data$best_model_overall$final_fit$finalModel$err.rate) %>% # Extract error rates
  mutate(ntrees = 1:nrow(best_class_data$best_model_overall$final_fit$finalModel$err.rate)) # Add number of trees

# Plot OOB error
ggplot(err_rate, aes(x = ntrees, y = OOB)) +
  geom_line(color = "skyblue1", linewidth = 1) +
  labs(x = "Number of Trees", y = "OOB Error", title = "Out Of Bag (OOB) Error") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) # Plot OOB error
```

# 5. Testing on the best model
```{r testing, echo=FALSE}
# Get predicted probabilities for the positive class ("will_die_within_10_years")
pred_probs <- predict(best_class_data$best_model_overall, newdata = matching_test_split, type = "prob")
positive_class_probs <- pred_probs[, "will_die_within_10_years"] # Select probabilities for "will_die_within_10_years"

# Generate the ROC curve for the test set
roc_curve <- roc(matching_test_split$prognosis, positive_class_probs, levels = levels(matching_test_split$prognosis), direction = ">")

# Create Precision-Recall curve for the test set
pr_curve <- pr.curve(
  scores.class0 = positive_class_probs, # Probabilities for the positive class
  weights.class0 = as.numeric(matching_test_split$prognosis == "will_die_within_10_years"), # Binary vector for positive class
  curve = TRUE # Return curve for plotting
)

# Save precision and recall plots as PNG
png("precision_recall_plots_train_test.png", width = 1600, height = 800) # Save output to PNG file

# Plot combined graphs (ROC and Precision-Recall curves side by side)
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 2) + 0.1) # Set layout to 1 row, 2 columns

# Plot ROC curve for both training and testing
plot(best_class_data$best_model_overall$roc, col = "red", main = "ROC", las = 1) # Training ROC
plot(roc_curve, col = "blue", add = TRUE, main = "ROC", las = 1) # Testing ROC
legend("bottomright", # Add legend with AUC values
  legend = c(
    paste0("AUC (Training) = ", signif(pROC::auc(best_class_data$best_model_overall$roc), 2)),
    paste0("AUC (Testing) = ", signif(roc_curve$auc, 2))
  ),
  col = c("red", "blue"), lty = 1, bty = "n"
)

# Plot Precision-Recall curve for both training and testing
plot(best_class_data$best_model_overall$prc, col = "red", main = "Precision-Recall Curve", xlab = "Recall", ylab = "Precision") # Training PRC
lines(pr_curve$curve[, 3], pr_curve$curve[, 1], col = "blue") # Testing PRC
legend("topright", # Add legend with AUC values
  legend = c(
    paste0("AUC (Training) = ", round(best_class_data$best_model_overall$prc$auc, 2)),
    paste0("AUC (Testing) = ", round(pr_curve$auc.integral, 2))
  ),
  col = c("red", "blue"),
  lty = 1,
  bty = "n"
)

# Reset plotting parameters and close PNG file
par(op)
dev.off()

# Find the best threshold from the ROC curve
best_threshold <- as.numeric(coords(roc_curve, "best", ret = "threshold")) # Threshold maximizing sensitivity-specificity
print(paste("Best threshold based on ROC curve:", best_threshold))

# Apply the best threshold for classification
pred_best_threshold <- ifelse(positive_class_probs > best_threshold, "will_die_within_10_years", "will_die_within_5_years") # Classify based on threshold

# Ensure predicted classes have the correct factor levels
pred_best_threshold <- factor(pred_best_threshold, levels = levels(matching_test_split$prognosis))

# Calculate accuracy on the test set using the best threshold
pred_table <- table(pred_best_threshold, matching_test_split$prognosis) # Create confusion matrix table
dimnames(pred_table) <- list( # Assign dimension names to match factor levels
  c("will_die_within_10_years", "will_die_within_5_years"),
  c("will_die_within_10_years", "will_die_within_5_years")
)
test_confusion <- confusionMatrix(pred_table) # Compute confusion matrix
test_confusion

# Plot the confusion matrix for the test set
confusionPlot(test_confusion, title = "Confusion Matrix (Testing Set)")
```